Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-10-13 ltx/paper.tex
added some results images to paper. left other parts of paper mostly alone.

\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

% \subsection{Related Work}

@@

Data gained by performing extraction on videos containing objects is significantly time-dependent. Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene. The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2006} in 2006.

@@

\begin{figure}
\centering
{\includegraphics[width=45mm]{antpic2.png}}
%\hspace{2mm}
{\includegraphics[width=60mm]{frame_diff_ants_1.png}}
\caption{(a.) Frame of a video containing five ants. (b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here... these aren't actually from the same frame, and also formatting is screwed up)}}
\label{test}
\end{figure}

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpudpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. The MCMC sampler described in this section operates in this alternate representation, which we can formulate as
\begin{align}
\begin{split}
	 \bold{x}_{i} | c_{i}, \theta_{c_{i}} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i}}, \Sigma_{c_{i}})  \mathcal{M}n(\bold{p}_{c_{i}}) \\
	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1 \\
	\theta_{k, t} | \theta_{k, t-1}   &\sim 
\begin{cases}
	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
	\mathbb{G}_{0}  \hspace{15pt} \text{if } k = K_{t-1} + 1
\end {cases}
\end{split}
\end{align}
where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}. This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in \willie{cite figure}.

@@

\subsection{Modeling Background Noise}
\willie{Do we think this is worth it? I can make extraction pretty clean on most datasets and this adds complication to an already complicated model}

Noisy extraction procedures may result in erroneous data points that do not correspond to any foreground object. We would like to incorporate these points into our model. To do this, we modify our base distribution $\mathbb{G}_{0}$

@@

@@

@@

@@

\subsubsection{PETS2001 or VS-PETS/Soccer}

\subsubsection{PETS2004/Caviar}

>>>>

\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\subsection{Related Work}

@@

Data gained by performing extraction on videos containing objects is significantly time-dependent. Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene. The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2007} in 2006.

@@

% \begin{figure}
% \centering{
% \includegraphics[width=45mm]{antpic2.png}}
% %\hspace{2mm}
% {\includegraphics[width=60mm]{frame_diff_ants_1.png}}
% \caption{(a.) Frame of a video containing five ants. (b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here... these aren't actually from the same frame, and also formatting is screwed up)}}
% \label{test}
% \end{figure}

\begin{figure}
  \centering               
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.5\textwidth]{antpic2.png}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.5\textwidth]{frame_diff_ants_1.png}}
  \caption{(a.) Frame of a video containing five ants. (b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here... these aren't actually from the same frame, and also formatting is screwed up)}}
  \label{fig:img_and_framediff}
\end{figure}

@@

The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpudpm}. This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions. To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time. Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster . At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables. The MCMC sampler described in this section operates in this alternate representation, which we can formulate as
\begin{align}
\begin{split}
	 \bold{x}_{i} | c_{i}, \theta_{c_{i}} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i}}, \Sigma_{c_{i}})  \mathcal{M}n(\bold{p}_{c_{i}}) \\
	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
	d_{i}-i-1 | \rho  &\sim \text{Geo}(\rho)\\
	\theta_{k, t} | \theta_{k, t-1} \willie{other-prior-params}  &\sim 
\begin{cases}
	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
	\mathbb{G}_{0}(\willie{prior params})  \hspace{15pt} \text{if } k = K_{t-1} + 1
\end {cases}
\end{split}
\end{align}
where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}. This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in \willie{cite figure}.

@@

\subsection{Modeling Background Noise}

Noisy extraction procedures may result in erroneous data points that do not correspond to any foreground object. We would like to incorporate these points into our model. To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution \willie{need more rigorus terms here} with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.5\textwidth]{cross_final}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{bounce_final}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though the objects reverse direction in the second video. \willie{will show assignments by marker type in addition to color}}
  \label{fig:synth_one_plot}
\end{figure}

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{synth2_img}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{synth2_final}}
  \caption{(a.) Video frame $f=10$ from synthetic experiment two. (b.) Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown. The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.  \willie{will show assignments by marker type in addition to color}}
  \label{fig:synth_two_plot}
\end{figure}

@@

%\subsubsection{PETS2000}

\subsubsection{PETS2001}

The PETS2001 dataset consists of

\begin{figure}
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_4}}
  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlayed. \willie{will show path trail over prev 10 frames}}
  \label{fig:pets2001_imgs}
\end{figure}

\begin{figure}
	\centering
	\subfloat[]{\includegraphics[width=0.52\textwidth]{pets2001_results_1}}
	\subfloat[]{\includegraphics[width=0.48\textwidth]{pets2001_results_2}}
	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown. \willie{will show assignments by marker type in addition to color}}
	\label{fig:pets2001_results}
\end{figure}

%\subsubsection{PETS2004/Caviar}

The PETS2009/2010 dataset consists of

\begin{figure}
  \centering             
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_1}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_2}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_3}}
  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_4}}
  \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlayed. \willie{will show path trail over prev 10 frames}}
  \label{fig:pets2009_imgs}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=3in]{pets2009_results}
	% \subfloat[]{\includegraphics[width=0.52\textwidth]{pets2001_results_1}}
	% \subfloat[]{\includegraphics[width=0.48\textwidth]{pets2001_results_2}}
	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown. \willie{will show assignments by marker type in addition to color}}
	\label{fig:pets2009_results}
\end{figure}

@@

The PETS2000 dataset consists of

\begin{figure}
	\centering
	\subfloat[]{\includegraphics[width=0.5\textwidth]{pets2000_results_1}}
	\subfloat[]{\includegraphics[width=0.5\textwidth]{pets2000_results_2}}
	\caption{\willie{This dataset has not done that well so far. Not sure why. Here are two typical cases.}}
	\label{fig:pets2000_results}
\end{figure}

\subsubsection{PETS2004/Caviar Datasets}

The Caviar datasets used in PETS2004 consist of multiple test-case-scenarios involving one or more humans.

\begin{figure}
	\centering
	\includegraphics[width=3in]{pets2004_results_1}
	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlayed on frame $f=20$.}
	\label{fig:pets2009_results}
\end{figure}
